---
title: "Healthcare Provider Anomaly Detection Dossier"
subtitle: "Zero-Dependency Algorithmic Fraud Detection Engine"
author: "Anomaly Detection Engine v1.0"
date: "`r format(Sys.time(), '%B %d, %Y — %H:%M')`"
output:
  html_document:
    theme: null
    highlight: pygments
    toc: true
    toc_depth: 3
    number_sections: true
    css: null
    self_contained: true
---

```{css, echo=FALSE}
/* ═══════════════════════════════════════════════════════════════════
   DARK THEME — ANOMALY DETECTION DOSSIER
   ═══════════════════════════════════════════════════════════════════ */

@import url('https://fonts.googleapis.com/css2?family=JetBrains+Mono:wght@400;700&family=Inter:wght@300;400;600;700;900&display=swap');

:root {
  --bg-primary: #0a0e17;
  --bg-secondary: #111827;
  --bg-card: #1a2036;
  --bg-card-hover: #212945;
  --border: #2a3555;
  --text-primary: #e2e8f0;
  --text-secondary: #94a3b8;
  --text-muted: #64748b;
  --accent-blue: #3b82f6;
  --accent-cyan: #06b6d4;
  --accent-purple: #8b5cf6;
  --accent-green: #10b981;
  --accent-yellow: #f59e0b;
  --accent-orange: #f97316;
  --accent-red: #ef4444;
  --accent-pink: #ec4899;
  --gradient-1: linear-gradient(135deg, #3b82f6, #8b5cf6);
  --gradient-2: linear-gradient(135deg, #06b6d4, #3b82f6);
  --gradient-danger: linear-gradient(135deg, #ef4444, #f97316);
}

* { box-sizing: border-box; }

body {
  background: var(--bg-primary);
  color: var(--text-primary);
  font-family: 'Inter', -apple-system, BlinkMacSystemFont, sans-serif;
  font-size: 15px;
  line-height: 1.7;
  margin: 0;
  padding: 0;
}

.main-container {
  max-width: 1200px;
  margin: 0 auto;
  padding: 2rem 2.5rem;
  background: var(--bg-primary);
}

/* ── HEADERS ── */
h1, h2, h3, h4, h5, h6 {
  font-family: 'Inter', sans-serif;
  color: var(--text-primary);
  font-weight: 700;
  letter-spacing: -0.02em;
  margin-top: 2.5rem;
}

h1.title {
  font-size: 2.5rem;
  font-weight: 900;
  background: var(--gradient-1);
  -webkit-background-clip: text;
  -webkit-text-fill-color: transparent;
  background-clip: text;
  margin-bottom: 0.25rem;
  letter-spacing: -0.03em;
}

h1.title + h3.subtitle {
  color: var(--text-secondary);
  font-weight: 400;
  font-size: 1.1rem;
  margin-top: 0;
}

h4.author, h4.date {
  color: var(--text-muted);
  font-weight: 400;
  font-size: 0.9rem;
  margin: 0.25rem 0;
}

h1 { font-size: 1.8rem; border-bottom: 2px solid var(--border); padding-bottom: 0.5rem; }
h2 { font-size: 1.4rem; color: var(--accent-cyan); }
h3 { font-size: 1.15rem; color: var(--accent-purple); }

/* ── CODE ── */
code {
  font-family: 'JetBrains Mono', monospace;
  font-size: 0.85em;
  background: var(--bg-secondary);
  color: var(--accent-cyan);
  padding: 0.15em 0.4em;
  border-radius: 4px;
  border: 1px solid var(--border);
}

pre {
  background: rgba(17, 24, 39, 0.7) !important;
  backdrop-filter: blur(10px);
  border: 1px solid rgba(255, 255, 255, 0.08);
  border-radius: 8px;
  padding: 1rem 1.25rem;
  overflow-x: auto;
  font-size: 0.85rem;
  box-shadow: inset 0 2px 4px rgba(0,0,0,0.5);
}

pre code {
  background: none;
  border: none;
  padding: 0;
  color: var(--text-primary);
}

/* ── TABLES ── */
table {
  width: 100%;
  border-collapse: collapse;
  margin: 1.5rem 0;
  font-size: 0.9rem;
  background: rgba(26, 32, 54, 0.4);
  backdrop-filter: blur(12px);
  border-radius: 8px;
  overflow: hidden;
  border: 1px solid rgba(255, 255, 255, 0.1);
  box-shadow: 0 8px 32px rgba(0,0,0,0.2);
}

thead th {
  background: var(--bg-secondary);
  color: var(--accent-cyan);
  font-weight: 600;
  text-transform: uppercase;
  font-size: 0.75rem;
  letter-spacing: 0.05em;
  padding: 0.75rem 1rem;
  border-bottom: 2px solid var(--accent-blue);
  text-align: left;
}

tbody td {
  padding: 0.6rem 1rem;
  border-bottom: 1px solid var(--border);
  color: var(--text-primary);
}

tbody tr:hover {
  background: var(--bg-card-hover);
}

tbody tr:last-child td { border-bottom: none; }

/* ── TOC ── */
.tocify {
  background: var(--bg-secondary) !important;
  border: 1px solid var(--border) !important;
  border-radius: 8px !important;
}

.tocify-header, .tocify-subheader {
  font-size: 0.85rem;
}

.tocify a {
  color: var(--text-secondary) !important;
  padding: 0.4rem 1rem !important;
}

.tocify .active a {
  color: var(--accent-cyan) !important;
  background: var(--bg-card) !important;
  border-left: 3px solid var(--accent-blue) !important;
}

/* ── PLOTS ── */
img, .plot-container {
  max-width: 100%;
  border-radius: 8px;
  margin: 1rem 0;
}

p img {
  border: 1px solid var(--border);
  border-radius: 8px;
}

/* ── BLOCKQUOTES (info boxes) ── */
blockquote {
  background: var(--bg-card);
  border-left: 4px solid var(--accent-blue);
  padding: 1rem 1.25rem;
  margin: 1.5rem 0;
  border-radius: 0 8px 8px 0;
  color: var(--text-secondary);
}

/* ── LINKS ── */
a { color: var(--accent-blue); text-decoration: none; }
a:hover { color: var(--accent-cyan); text-decoration: underline; }

/* ── HORIZONTAL RULES ── */
hr {
  border: none;
  height: 1px;
  background: var(--border);
  margin: 2rem 0;
}

/* ── CUSTOM CLASSES ── */
.metric-box {
  display: inline-block;
  background: rgba(26, 32, 54, 0.6);
  backdrop-filter: blur(10px);
  border: 1px solid rgba(255, 255, 255, 0.1);
  box-shadow: 0 8px 32px rgba(0, 0, 0, 0.3);
  border-radius: 12px;
  padding: 1rem 1.5rem;
  margin: 0.5rem;
  text-align: center;
  min-width: 160px;
  transition: transform 0.2s ease, box-shadow 0.2s ease;
}

.metric-box:hover {
  transform: translateY(-5px);
  box-shadow: 0 12px 40px rgba(0, 0, 0, 0.5);
}

.metric-value {
  font-size: 2rem;
  font-weight: 900;
  font-family: 'JetBrains Mono', monospace;
}

.metric-label {
  font-size: 0.75rem;
  text-transform: uppercase;
  letter-spacing: 0.05em;
  color: var(--text-muted);
  margin-top: 0.25rem;
}

.risk-critical { color: var(--accent-red); }
.risk-high { color: var(--accent-orange); }
.risk-elevated { color: var(--accent-yellow); }
.risk-moderate { color: var(--accent-blue); }
.risk-low { color: var(--accent-green); }

/* ── SCROLLBAR ── */
::-webkit-scrollbar { width: 8px; height: 8px; }
::-webkit-scrollbar-track { background: var(--bg-primary); }
::-webkit-scrollbar-thumb { background: var(--border); border-radius: 4px; }
::-webkit-scrollbar-thumb:hover { background: var(--text-muted); }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = FALSE,
  warning    = FALSE,
  message    = FALSE,
  fig.width  = 10,
  fig.height = 6,
  fig.retina = 2,
  dpi        = 150
)

library(ggplot2)
library(knitr)

# ── Dark ggplot2 theme ──
theme_anomaly <- function(base_size = 13) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      plot.background   = element_rect(fill = "#0a0e17", color = NA),
      panel.background  = element_rect(fill = "#111827", color = NA),
      panel.grid.major  = element_line(color = "#1e293b", linewidth = 0.3),
      panel.grid.minor  = element_blank(),
      text              = element_text(color = "#e2e8f0", family = "sans"),
      plot.title        = element_text(color = "#e2e8f0", size = 16, face = "bold",
                                       hjust = 0, margin = margin(b = 8)),
      plot.subtitle     = element_text(color = "#94a3b8", size = 11,
                                       hjust = 0, margin = margin(b = 15)),
      axis.text         = element_text(color = "#94a3b8", size = 10),
      axis.title        = element_text(color = "#cbd5e1", size = 11, face = "bold"),
      legend.background = element_rect(fill = "#111827", color = NA),
      legend.text       = element_text(color = "#94a3b8"),
      legend.title      = element_text(color = "#e2e8f0", face = "bold"),
      legend.key        = element_rect(fill = "#111827", color = NA),
      strip.text        = element_text(color = "#06b6d4", face = "bold", size = 11),
      plot.margin       = margin(15, 15, 15, 15)
    )
}
theme_set(theme_anomaly())

# ── Color palettes ──
pal_risk <- c("CRITICAL" = "#ef4444", "HIGH" = "#f97316",
              "ELEVATED" = "#f59e0b", "MODERATE" = "#3b82f6", "LOW" = "#10b981")

pal_gradient <- c("#10b981", "#3b82f6", "#8b5cf6", "#f59e0b", "#ef4444")

pal_binary <- c("Normal" = "#3b82f6", "Anomaly" = "#ef4444")
```

```{r load-engines}
# ── Source all engine files ──
source("R/05_utils.R",           local = TRUE)
source("R/01_generate_data.R",   local = TRUE)
source("R/02_mvg_engine.R",      local = TRUE)
source("R/03_network_analysis.R",local = TRUE)
source("R/04_benford_law.R",     local = TRUE)
```

```{r run-engines, results='hide'}
# ══════════════════════════════════════════════════════════════════
# EXECUTE THE FULL DETECTION PIPELINE
# ══════════════════════════════════════════════════════════════════

# 1. Generate synthetic data
data <- generate_all_data(n_doctors = 500, seed = 42)
doctors      <- data$doctors
network      <- data$network
transactions <- data$transactions

# 2. Run MVG Anomaly Detection
mvg <- run_mvg_anomaly_detection(doctors)

# 3. Run Network Analysis
net <- run_network_analysis(doctors, network)

# 4. Run Benford's Law Analysis
ben <- run_benford_analysis(doctors, transactions)

# 5. Merge all results
results <- merge(doctors, mvg$results, by = "doctor_id", all.x = TRUE)
results <- merge(results, net$results, by = "doctor_id", all.x = TRUE)
results <- merge(results, ben$results, by = "doctor_id", all.x = TRUE)

# Fill NA scores with neutral 0.5
results$mvg_anomaly_score[is.na(results$mvg_anomaly_score)] <- 0
results$network_anomaly_score[is.na(results$network_anomaly_score)] <- 0
results$benford_score[is.na(results$benford_score)] <- 0

# 6. Compute composite risk score
results$composite_score <- composite_risk_score(
  results$mvg_anomaly_score,
  results$network_anomaly_score,
  results$benford_score
)

# 7. Classify risk
results$risk_level <- classify_risk(results$composite_score)

# 8. Sort by risk
results <- results[order(-results$composite_score), ]
```

---

# Executive Summary

```{r executive-summary}
n_total     <- nrow(results)
n_flagged   <- sum(results$composite_score >= 0.5)
n_critical  <- sum(results$risk_level == "CRITICAL")
n_high      <- sum(results$risk_level == "HIGH")
n_true_anom <- sum(results$is_anomaly)
detected    <- sum(results$is_anomaly & results$composite_score >= 0.5)
precision   <- sum(results$is_anomaly & results$composite_score >= 0.5) / max(1, n_flagged)
recall      <- detected / max(1, n_true_anom)
```

<div style="text-align: center; margin: 2rem 0;">
<div class="metric-box">
<div class="metric-value" style="color: #06b6d4;">`r n_total`</div>
<div class="metric-label">Providers Analyzed</div>
</div>
<div class="metric-box">
<div class="metric-value risk-critical">`r n_flagged`</div>
<div class="metric-label">Flagged Suspicious</div>
</div>
<div class="metric-box">
<div class="metric-value risk-critical">`r n_critical`</div>
<div class="metric-label">Critical Risk</div>
</div>
<div class="metric-box">
<div class="metric-value" style="color: #f97316;">`r n_high`</div>
<div class="metric-label">High Risk</div>
</div>
<div class="metric-box">
<div class="metric-value" style="color: #10b981;">`r sprintf("%.0f%%", recall * 100)`</div>
<div class="metric-label">Detection Recall</div>
</div>
<div class="metric-box">
<div class="metric-value" style="color: #8b5cf6;">`r sprintf("%.0f%%", precision * 100)`</div>
<div class="metric-label">Precision</div>
</div>
</div>

> **Engine Summary:** The anomaly detection pipeline analyzed **`r n_total`** healthcare provider profiles
> using three independent mathematical methods — Multivariate Gaussian probability density,
> Eigenvector Centrality network analysis, and Benford's Law digit forensics — producing a
> composite fraud risk score for each provider. **`r n_flagged`** providers were flagged as
> suspicious (composite score ≥ 0.50), with **`r n_critical`** classified as **CRITICAL** risk.

---

# Methodology

## Mathematical Framework

This engine employs three orthogonal detection methods, each implemented using scratch-built linear algebra:

### Multivariate Gaussian Probability Density

Each provider is modeled as a point in $\mathbb{R}^7$ feature space. The probability of observing that provider under the estimated normal distribution is:

$$p(\mathbf{x}; \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{1}{(2\pi)^{n/2}|\boldsymbol{\Sigma}|^{1/2}} \exp\left(-\frac{1}{2}(\mathbf{x}-\boldsymbol{\mu})^T\boldsymbol{\Sigma}^{-1}(\mathbf{x}-\boldsymbol{\mu})\right)$$

The covariance matrix $\boldsymbol{\Sigma}$ is computed from scratch: $\boldsymbol{\Sigma} = \frac{1}{m-1}(\mathbf{X} - \boldsymbol{\mu})^T(\mathbf{X} - \boldsymbol{\mu})$

### Eigenvector Centrality (Power Iteration)

The referral network adjacency matrix $\mathbf{A}$ is analyzed via iterative matrix multiplication:

$$\mathbf{v}_{k+1} = \frac{\mathbf{A}\mathbf{v}_k}{||\mathbf{A}\mathbf{v}_k||}$$

Convergence yields the dominant eigenvector, whose components give each node's centrality score — identifying hubs of suspicious referral activity.

### Benford's Law

Leading-digit distribution of billing amounts is tested against the theoretical Benford distribution:

$$P(d) = \log_{10}\left(1 + \frac{1}{d}\right), \quad d = 1, 2, \ldots, 9$$

Deviations are quantified via $\chi^2$ goodness-of-fit and Mean Absolute Deviation (MAD).

---

# Multivariate Gaussian Analysis {.tabset}

## Mahalanobis Distance Scatter

```{r mahal-scatter, fig.height=7}
plot_data <- results
plot_data$label <- ifelse(results$is_anomaly, "Anomaly", "Normal")
plot_data$alpha_val <- ifelse(results$is_anomaly, 1, 0.5)

ggplot(plot_data, aes(x = seq_len(nrow(plot_data)), y = mahal_distance)) +
  geom_point(aes(color = label, size = mvg_anomaly_score), alpha = 0.7) +
  geom_hline(yintercept = quantile(results$mahal_distance, 0.90, na.rm = TRUE),
             color = "#ef4444", linetype = "dashed", linewidth = 0.8) +
  annotate("text", x = nrow(plot_data) * 0.85,
           y = quantile(results$mahal_distance, 0.90, na.rm = TRUE) + 0.5,
           label = "90th Percentile Threshold", color = "#ef4444",
           size = 3.5, fontface = "bold") +
  scale_color_manual(values = pal_binary) +
  scale_size_continuous(range = c(1, 5), guide = "none") +
  labs(
    title = "Mahalanobis Distance — All Healthcare Providers",
    subtitle = "Each point represents a provider. Distance from the statistical center reveals anomalous profiles.",
    x = "Provider Index",
    y = "Mahalanobis Distance",
    color = "Ground Truth"
  )
```

## Probability Density Distribution

```{r prob-density, fig.height=6}
plot_data <- results
plot_data$label <- ifelse(results$is_anomaly, "Anomaly", "Normal")
plot_data$log_prob <- log10(pmax(plot_data$probability, 1e-300))

ggplot(plot_data, aes(x = log_prob, fill = label)) +
  geom_histogram(bins = 50, alpha = 0.75, position = "identity", color = "#0a0e17") +
  geom_vline(xintercept = log10(mvg$epsilon), color = "#ef4444",
             linetype = "dashed", linewidth = 0.8) +
  annotate("text", x = log10(mvg$epsilon) - 1.5, y = Inf,
           label = paste("ε =", format_probability(mvg$epsilon)),
           color = "#ef4444", size = 3.5, fontface = "bold", vjust = 2) +
  scale_fill_manual(values = pal_binary) +
  labs(
    title = "Log-Probability Distribution Under Multivariate Gaussian Model",
    subtitle = "Providers left of the threshold ε have statistically improbable feature combinations",
    x = expression(log[10](p(x))),
    y = "Count",
    fill = "Ground Truth"
  )
```

## Feature Correlation Heatmap

```{r covariance-heatmap, fig.height=7}
# Correlation matrix from the covariance matrix
Sigma <- mvg$Sigma
D_inv <- diag(1 / sqrt(diag(Sigma)))
cor_mat <- D_inv %*% Sigma %*% D_inv
colnames(cor_mat) <- mvg$feature_cols
rownames(cor_mat) <- mvg$feature_cols

# Reshape for ggplot2
cor_df <- expand.grid(Var1 = mvg$feature_cols, Var2 = mvg$feature_cols)
cor_df$value <- as.vector(cor_mat)

ggplot(cor_df, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "#1e293b", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.2f", value)), color = "#e2e8f0", size = 3) +
  scale_fill_gradient2(low = "#3b82f6", mid = "#111827", high = "#ef4444",
                       midpoint = 0, limits = c(-1, 1)) +
  labs(
    title = "Feature Correlation Matrix (Derived from Scratch-Built Σ)",
    subtitle = "Computed via (X−μ)ᵀ(X−μ) / (m−1) — no cov() function used",
    x = NULL, y = NULL, fill = "Correlation"
  ) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1, size = 9),
    axis.text.y = element_text(size = 9)
  )
```

## Anomaly Score vs Features

```{r anomaly-feature-panels, fig.height=8}
plot_df <- data.frame(
  score = results$mvg_anomaly_score,
  billing = results$avg_billing_amount,
  consult = results$consultations_per_month,
  complaints = results$patient_complaints,
  prescriptions = results$prescriptions_per_patient,
  label = ifelse(results$is_anomaly, "Anomaly", "Normal")
)

p1 <- ggplot(plot_df, aes(x = billing, y = score, color = label)) +
  geom_point(alpha = 0.6, size = 1.5) +
  scale_color_manual(values = pal_binary) +
  labs(x = "Avg Billing Amount", y = "MVG Score", color = NULL) +
  theme(legend.position = "none")

p2 <- ggplot(plot_df, aes(x = consult, y = score, color = label)) +
  geom_point(alpha = 0.6, size = 1.5) +
  scale_color_manual(values = pal_binary) +
  labs(x = "Consultations/Month", y = "MVG Score", color = NULL) +
  theme(legend.position = "none")

p3 <- ggplot(plot_df, aes(x = complaints, y = score, color = label)) +
  geom_point(alpha = 0.6, size = 1.5) +
  scale_color_manual(values = pal_binary) +
  labs(x = "Patient Complaints", y = "MVG Score", color = NULL) +
  theme(legend.position = "none")

p4 <- ggplot(plot_df, aes(x = prescriptions, y = score, color = label)) +
  geom_point(alpha = 0.6, size = 1.5) +
  scale_color_manual(values = pal_binary) +
  labs(x = "Prescriptions/Patient", y = "MVG Score", color = NULL)

gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2,
                        top = grid::textGrob(
                          "MVG Anomaly Score vs Key Features",
                          gp = grid::gpar(fontsize = 15, fontface = "bold", col = "#e2e8f0")))
```

---

# Network Analysis: Collusion Detection {.tabset}

## Eigenvector Centrality Distribution

```{r centrality-dist, fig.height=6}
net_df <- results[!is.na(results$eigenvector_centrality), ]
net_df$label <- ifelse(net_df$is_anomaly, "Anomaly", "Normal")

ggplot(net_df, aes(x = eigenvector_centrality, fill = label)) +
  geom_histogram(bins = 40, alpha = 0.75, position = "identity", color = "#0a0e17") +
  geom_vline(xintercept = net$collusion$threshold, color = "#f97316",
             linetype = "dashed", linewidth = 0.8) +
  annotate("text", x = net$collusion$threshold, y = Inf,
           label = "Hub Threshold  ", color = "#f97316",
           size = 3.5, fontface = "bold", vjust = 1.5, hjust = 1, angle = 90) +
  scale_fill_manual(values = pal_binary) +
  labs(
    title = "Eigenvector Centrality Distribution (Power Iteration)",
    subtitle = paste("Converged in", net$eig_result$iterations,
                     "iterations | Dominant eigenvalue λ =",
                     round(net$eig_result$eigenvalue, 2)),
    x = "Eigenvector Centrality Score",
    y = "Count",
    fill = "Ground Truth"
  )
```

## Centrality vs Anomaly Score

```{r centrality-vs-mvg, fig.height=6}
net_df <- results[!is.na(results$eigenvector_centrality), ]
net_df$label <- ifelse(net_df$is_anomaly, "Anomaly", "Normal")

ggplot(net_df, aes(x = eigenvector_centrality, y = mvg_anomaly_score,
                   color = label, size = composite_score)) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = pal_binary) +
  scale_size_continuous(range = c(1, 6), name = "Composite\nScore") +
  labs(
    title = "Network Centrality vs Statistical Anomaly Score",
    subtitle = "Providers in the upper-right quadrant are both statistically anomalous AND network hubs",
    x = "Eigenvector Centrality",
    y = "MVG Anomaly Score",
    color = "Ground Truth"
  )
```

## Community Structure

```{r community-plot, fig.height=6}
net_df <- results[!is.na(results$community) & results$community > 0, ]
net_df$community_label <- paste("Community", net_df$community)
net_df$label <- ifelse(net_df$is_anomaly, "Anomaly", "Normal")

comm_summary <- aggregate(is_anomaly ~ community, data = net_df, FUN = mean)
comm_summary$pct_label <- paste0(round(comm_summary$is_anomaly * 100, 1), "% anomalous")

ggplot(net_df, aes(x = factor(community), fill = label)) +
  geom_bar(position = "stack", alpha = 0.85, color = "#0a0e17") +
  scale_fill_manual(values = pal_binary) +
  labs(
    title = "Spectral Community Composition",
    subtitle = "Detected via Laplacian eigendecomposition (L = D − A) with scratch k-means clustering",
    x = "Community ID",
    y = "Number of Providers",
    fill = "Ground Truth"
  )
```

---

# Benford's Law Analysis {.tabset}

## Global Digit Distribution

```{r benford-global, fig.height=6}
expected <- benford_expected()
observed <- ben$global_observed$proportions

benford_df <- data.frame(
  digit = rep(1:9, 2),
  proportion = c(observed, expected),
  type = rep(c("Observed", "Expected (Benford)"), each = 9)
)

ggplot(benford_df, aes(x = factor(digit), y = proportion, fill = type)) +
  geom_col(position = "dodge", alpha = 0.85, color = "#0a0e17", width = 0.7) +
  scale_fill_manual(values = c("Observed" = "#3b82f6", "Expected (Benford)" = "#8b5cf6")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Global Leading-Digit Distribution vs Benford's Law",
    subtitle = sprintf("χ² = %.2f (p = %s) | MAD = %.4f — %s",
                       ben$global_chi$chi_squared,
                       format_probability(ben$global_chi$p_value),
                       ben$global_mad$mad,
                       ben$global_mad$classification),
    x = "Leading Digit",
    y = "Proportion",
    fill = NULL
  )
```

## Per-Doctor Benford Conformity

```{r benford-per-doctor, fig.height=6}
ben_df <- results[!is.na(results$benford_mad), ]
ben_df$label <- ifelse(ben_df$is_anomaly, "Anomaly", "Normal")

ggplot(ben_df, aes(x = benford_mad, fill = label)) +
  geom_histogram(bins = 40, alpha = 0.75, position = "identity", color = "#0a0e17") +
  geom_vline(xintercept = 0.006, color = "#10b981", linetype = "dotted", linewidth = 0.6) +
  geom_vline(xintercept = 0.012, color = "#f59e0b", linetype = "dotted", linewidth = 0.6) +
  geom_vline(xintercept = 0.015, color = "#ef4444", linetype = "dashed", linewidth = 0.8) +
  annotate("text", x = 0.006, y = Inf, label = "Close  ", color = "#10b981",
           size = 3.5, fontface = "bold", hjust = 1, vjust = -0.5, angle = 90) +
  annotate("text", x = 0.012, y = Inf, label = "Acceptable  ", color = "#f59e0b",
           size = 3.5, fontface = "bold", hjust = 1, vjust = -0.5, angle = 90) +
  annotate("text", x = 0.015, y = Inf, label = "Nonconforming  ", color = "#ef4444",
           size = 3.5, fontface = "bold", hjust = 1, vjust = -0.5, angle = 90) +
  scale_fill_manual(values = pal_binary) +
  labs(
    title = "Per-Doctor Benford's Law Mean Absolute Deviation (MAD)",
    subtitle = "Nigrini's conformity thresholds shown. Higher MAD = more likely fabricated billing data.",
    x = "Mean Absolute Deviation from Benford's Law",
    y = "Count",
    fill = "Ground Truth"
  )
```

## Benford Classification Breakdown

```{r benford-classification, fig.height=5}
class_df <- as.data.frame(table(results$benford_class, results$is_anomaly))
colnames(class_df) <- c("Classification", "IsAnomaly", "Count")
class_df$IsAnomaly <- ifelse(class_df$IsAnomaly == TRUE, "Anomaly", "Normal")

class_order <- c("CLOSE CONFORMITY", "ACCEPTABLE CONFORMITY",
                 "MARGINALLY ACCEPTABLE", "NONCONFORMING", "INSUFFICIENT DATA")
class_df$Classification <- factor(class_df$Classification, levels = class_order)

ggplot(class_df, aes(x = Classification, y = Count, fill = IsAnomaly)) +
  geom_col(position = "stack", alpha = 0.85, color = "#0a0e17") +
  scale_fill_manual(values = pal_binary) +
  labs(
    title = "Benford Conformity Classification",
    subtitle = "Anomalous providers disproportionately appear in non-conforming categories",
    x = NULL, y = "Number of Providers", fill = "Ground Truth"
  ) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

---

# Composite Risk Assessment {.tabset}

## Risk Distribution

```{r risk-dist, fig.height=6}
risk_df <- as.data.frame(table(results$risk_level))
colnames(risk_df) <- c("Risk", "Count")
risk_df$Risk <- factor(risk_df$Risk, levels = c("CRITICAL", "HIGH", "ELEVATED", "MODERATE", "LOW"))

ggplot(risk_df, aes(x = Risk, y = Count, fill = Risk)) +
  geom_col(alpha = 0.9, color = "#0a0e17", width = 0.7) +
  geom_text(aes(label = Count), vjust = -0.5, color = "#e2e8f0",
            fontface = "bold", size = 4.5) +
  scale_fill_manual(values = pal_risk) +
  labs(
    title = "Composite Risk Level Distribution",
    subtitle = "Weighted fusion: 50% MVG + 30% Network + 20% Benford",
    x = NULL, y = "Number of Providers"
  ) +
  theme(legend.position = "none")
```

## Composite Score vs Mahalanobis Distance

```{r composite-scatter, fig.height=7}
plot_df <- results
plot_df$label <- ifelse(results$is_anomaly, "Anomaly", "Normal")

ggplot(plot_df, aes(x = mahal_distance, y = composite_score,
                    color = as.character(risk_level))) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_manual(values = pal_risk, name = "Risk Level") +
  geom_hline(yintercept = 0.5, color = "#ef4444", linetype = "dashed", linewidth = 0.6) +
  labs(
    title = "Composite Fraud Score vs Mahalanobis Distance",
    subtitle = "The dashed line marks the investigation threshold (score ≥ 0.50)",
    x = "Mahalanobis Distance from Population Center",
    y = "Composite Fraud Risk Score"
  )
```

## Three-Method Comparison

```{r three-method, fig.height=7}
comp_df <- results[!is.na(results$network_anomaly_score) & !is.na(results$benford_score), ]
comp_df$label <- ifelse(comp_df$is_anomaly, "Anomaly", "Normal")

ggplot(comp_df, aes(x = mvg_anomaly_score, y = network_anomaly_score,
                    color = benford_score, shape = label)) +
  geom_point(alpha = 0.7, size = 2.5) +
  scale_color_gradient2(low = "#10b981", mid = "#f59e0b", high = "#ef4444",
                        midpoint = 0.5, name = "Benford\nScore") +
  scale_shape_manual(values = c("Normal" = 16, "Anomaly" = 17)) +
  labs(
    title = "Three-Dimensional Fraud Signal Comparison",
    subtitle = "MVG (x-axis) × Network (y-axis) × Benford (color) — anomalies cluster in upper-right-red",
    x = "MVG Anomaly Score (Statistical)",
    y = "Network Anomaly Score (Centrality)",
    shape = "Ground Truth"
  )
```

---

# Top Flagged Providers

```{r top-flagged}
top_n <- min(25, nrow(results))
top <- head(results, top_n)

display_df <- data.frame(
  Rank         = 1:top_n,
  Doctor_ID    = top$doctor_id,
  Specialty    = top$specialty,
  Risk         = as.character(top$risk_level),
  Composite    = sprintf("%.3f", top$composite_score),
  MVG_Score    = sprintf("%.3f", top$mvg_anomaly_score),
  Network      = sprintf("%.3f", top$network_anomaly_score),
  Benford      = sprintf("%.3f", top$benford_score),
  Mahalanobis  = sprintf("%.2f", top$mahal_distance),
  Probability  = format_probability(top$probability),
  True_Anomaly = ifelse(top$is_anomaly, "✓ YES", "—")
)

kable(display_df, align = c("r", "l", "l", "l", rep("r", 6), "c"),
      caption = "Top 25 Most Suspicious Healthcare Providers — Ranked by Composite Fraud Score")
```

---

# Detection Performance

```{r confusion-matrix}
threshold <- 0.5
predicted_pos <- results$composite_score >= threshold
actual_pos    <- results$is_anomaly

tp <- sum(predicted_pos & actual_pos)
fp <- sum(predicted_pos & !actual_pos)
fn <- sum(!predicted_pos & actual_pos)
tn <- sum(!predicted_pos & !actual_pos)

accuracy  <- (tp + tn) / nrow(results)
precision <- tp / max(1, tp + fp)
recall    <- tp / max(1, tp + fn)
f1        <- 2 * precision * recall / max(1e-10, precision + recall)
```

<div style="text-align: center; margin: 2rem 0;">
<div class="metric-box">
<div class="metric-value" style="color: #10b981;">`r sprintf("%.1f%%", accuracy * 100)`</div>
<div class="metric-label">Accuracy</div>
</div>
<div class="metric-box">
<div class="metric-value" style="color: #8b5cf6;">`r sprintf("%.1f%%", precision * 100)`</div>
<div class="metric-label">Precision</div>
</div>
<div class="metric-box">
<div class="metric-value" style="color: #06b6d4;">`r sprintf("%.1f%%", recall * 100)`</div>
<div class="metric-label">Recall</div>
</div>
<div class="metric-box">
<div class="metric-value" style="color: #f59e0b;">`r sprintf("%.1f%%", f1 * 100)`</div>
<div class="metric-label">F1 Score</div>
</div>
</div>

### Confusion Matrix

```{r confusion-table}
conf_df <- data.frame(
  ` ` = c("**Predicted Positive**", "**Predicted Negative**"),
  `Actual Positive` = c(tp, fn),
  `Actual Negative` = c(fp, tn),
  check.names = FALSE
)
kable(conf_df, align = c("l", "r", "r"),
      caption = paste("Confusion Matrix at threshold =", threshold))
```

> **Interpretation:** The engine correctly identified **`r tp`** of **`r tp + fn`** actual anomalies
> (recall = `r sprintf("%.1f%%", recall * 100)`). The precision of `r sprintf("%.1f%%", precision * 100)` means
> that `r sprintf("%.0f%%", precision * 100)` of flagged providers were truly anomalous — a strong result
> for an unsupervised detection system.

---

# Technical Appendix {.tabset}

## Eigenvalue Spectrum

```{r eigenvalue-spectrum, fig.height=5}
eig_df <- data.frame(
  index = seq_along(net$laplacian_eigenvalues),
  eigenvalue = net$laplacian_eigenvalues
)

ggplot(eig_df, aes(x = index, y = eigenvalue)) +
  geom_line(color = "#06b6d4", linewidth = 1) +
  geom_point(color = "#3b82f6", size = 3) +
  labs(
    title = "Laplacian Eigenvalue Spectrum",
    subtitle = "Gaps indicate natural community boundaries (spectral clustering)",
    x = "Eigenvalue Index",
    y = "Eigenvalue"
  )
```

## Algorithm Summary

| Method | Implementation | Key Math |
|--------|---------------|----------|
| **MVG PDF** | Scratch-built via `%*%`, `solve()`, `det()` | $p(x) = \frac{1}{(2\pi)^{n/2}\|\Sigma\|^{1/2}} \exp(-\frac{1}{2}(x-\mu)^T\Sigma^{-1}(x-\mu))$ |
| **Covariance Matrix** | Raw matrix multiplication | $\Sigma = \frac{1}{m-1}(X-\mu)^T(X-\mu)$ |
| **Mahalanobis Distance** | Vectorized `%*%` + `rowSums` | $D_M = \sqrt{(x-\mu)^T\Sigma^{-1}(x-\mu)}$ |
| **Eigenvector Centrality** | Power iteration | $v_{k+1} = Av_k / \|Av_k\|$ |
| **Spectral Clustering** | Laplacian eigendecomposition | $L = D - A$, cluster on Fiedler vector |
| **K-Means** | Scratch iterative assignment | Centroid update via mean of assigned points |
| **Benford's Law** | Leading-digit extraction | $P(d) = \log_{10}(1 + 1/d)$ |
| **Chi-Squared Test** | Scratch Pearson's test | $\chi^2 = \sum (O_i - E_i)^2 / E_i$ |
| **MAD Conformity** | Nigrini's classification | $MAD = \frac{1}{9}\sum\|O_i - E_i\|$ |

## Covariance Matrix

```{r cov-matrix-display}
cov_df <- as.data.frame(round(mvg$Sigma, 4))
colnames(cov_df) <- mvg$feature_cols
rownames(cov_df) <- mvg$feature_cols
kable(cov_df, caption = "Computed Covariance Matrix Σ (computed from scratch via matrix multiplication)")
```

---

<div style="text-align: center; color: #64748b; margin-top: 3rem; padding: 2rem; border-top: 1px solid #2a3555;">
<p style="font-size: 1.1rem; font-weight: 600; color: #94a3b8;">
Healthcare Provider Anomaly Detection Engine v1.0</p>
<p>Zero external ML dependencies · Scratch-built linear algebra · Pure R</p>
<p style="font-size: 0.8rem; letter-spacing: 0.05em; text-transform: uppercase;">Project made by WuXin</p>
</div>
